---
title: "HW1"
author: "Elyse Zhang"
date: "6/21/2018"
output: html_document
---

##1 Assignment description and preparation for test
In preparation for the test below calculate different outputs of glm() using lm().

###1.1 Reading data
Use the data from file Week1_Homework_Project_Data.csv to estimate linear model with lm().
Analize summary() of the estimated linear model.

```{r}
dp = "/Users/Elyse/Documents/UChicago/Courses/Summer 2018/Linear Nonlinear/Course Material/Lecture 1/"
test_dat <- read.table(paste(dp,'Week1_Test_Sample.csv',sep = '/'), header=TRUE)
head(test_dat,10)

```

```{r}
lm.fit = lm(Output ~., data = test_dat)
summary(lm.fit)
```

**What can you tell about the data and the fit?**
**input 2 is not that important, and input 1 with output is a good fit**

###1.2 Reproducing outputs of glm() using lm()
By using any outputs of lm() or summary(lm.fit) calculate the following outputs of glm().

**You can use glm() to check your answers, but, please, do not use glm object or any functions applied to glm object to calculate your results.**

Calculate variables:

1. coefficients (5%)
2. residuals (5%)
3. fitted.values (5%)
4. linear.predictors (10%)
5. deviance (25%)
   (1). Use deviance() (10%)
   (2). Calculate deviance manually based on the definition given in the lecture (15%)
6. Akaike Information Criterion aic (25%)
   (1). Obtain it by using AIC() (10%)
   (2). Calculate it manually using the definition given in the lecture. (15%)
7. y (5%)
8. null.deviance (10%)
9. dispersion (10%)


```{r}
names(lm.fit)
names(summary(lm.fit))
```

```{r}
glm.fit = glm(Output~.,family=gaussian(link="identity"),data=test_dat)
summary(glm.fit)

names(glm.fit)
names(summary(glm.fit))
```

#### 1. Coefficients
```{r}
cbind(coef.lm = lm.fit$coefficients,coef.glm =glm.fit$coefficients)
coefficients = lm.fit$coefficients
```

2. Residuals.
```{r}
cbind(lm.residuals = lm.fit$residuals,lm.residuals = glm.fit$residuals)[1:10,]
residuals = lm.fit$residuals
```

3. fitted values
```{r}
fitted.values = lm.fit$fitted.values
```

4. Linear Predictors
```{r}
linear.predictors = lm.fit$fitted.values
```

5. Deviance
```{r}
(deviance = sum(lm.fit$residuals^2))
```

6.AIC
```{r}
Manual.Log.Likelihood<- function(lm.fit) {
  n = length(lm.fit$fitted.values)
  sigma = summary(lm.fit)$sigma
  return(-n*log(2*pi*sigma^2)/2-sum((lm.fit$residuals)^2)/(2*(sigma^2)))
}

(Log.Likelihood = Manual.Log.Likelihood(lm.fit))
(AIC.Manual = -2 * Log.Likelihood +2*(lm.fit$rank+1))

aic = AIC.Manual
```

7.y
```{r}
(sum(abs(test_dat[,1]-glm.fit$y)))
y = test_dat[,1]
```

8.null.deviance
```{r}
lm.null.fit = lm(Output ~ 1 ,data=test_dat)
(null.deviance = sum(lm.null.fit$residuals^2))
(glm.fit$null.deviance)
abs(null.deviance- glm.fit$null.deviance)
```


9.dispersion
```{r}
dispersion=summary(lm.fit)$sigma^2
```


```{r}
res<- list(Linear.Model.Data.lm = lm.fit,
              coefficients = coefficients,
              residuals = residuals,
              fitted.values = fitted.values,
              linear.predictors = linear.predictors,
              deviance = deviance,
              aic = aic,
              y = y,
              null.deviance = null.deviance,
              dispersion = dispersion
              )
```

```{r}
saveRDS(res, file = paste(dp,'result.rds',sep = '/'))
```

