---
title: "LNL Assignment 4"
author: "Elyse Zhang"
date: "7/19/2018"
output: html_document
---
### 1. Logistic regression problem statement
  1. Select an area of application, for example, biomedical studies, engineering, marketing, etc.
  2. For the selected area of application postulate a problem that could be solved using binomial regression. Think of the two versions of the experiment design: with prospective and retrospective sampling. Try to create the problem with two predictors.
  4. Find or simulate data for the model.
  5. Analize the use of the three different link functions and compare them to each other.
  6. Conduct the fit of binomial regression model using glm() with family=binomial.
  7. Describe the problem, all steps of obtaining solution and conclusions about the results.

### 2. Test
```{r}
dataPath<-"~/Documents/UChicago/Courses/Summer 2018/Linear Nonlinear/Course Material/Lecture 4/nonlinear_models_04_data"
```


```{r}
train_dat <- read.table(paste(dataPath,'Week4_Test_Sample_Train.csv',sep = '/'), header=TRUE)
test_dat <- read.table(paste(dataPath,'Week4_Test_Sample_Test.csv',sep = '/'), header=TRUE)

str(train_dat)
```

```{r}
glm.logit<-glm(Output~Predictor1+Predictor2,family=binomial(link="logit"),train_dat)
summary(glm.logit)
```

```{r}
glm.probit<-glm(Output~Predictor1+Predictor2,family=binomial(link="probit"),train_dat)
summary(glm.probit)
```




```{r}
Predicted.glm.logit = predict.glm(glm.logit, newdata = test_dat, type = 'response')
Predicted.Output.logit = as.integer(Predicted.glm.logit>=0.5)

#Predicted.glm.logit
#Predicted.Output.logit
```

```{r}
Predicted.glm.probit = predict.glm(glm.probit, newdata = test_dat, type = 'response')
Predicted.Output.probit = as.integer(Predicted.glm.probit>=0.5)
```

```{r}
x= cbind.data.frame(Predicted.Output.logit=Predicted.Output.logit,
            Predicted.Output.probit=Predicted.Output.probit)

sum(x$Predicted.Output.logit == x$Predicted.Output.probit)
identical(x$Predicted.Output.logit, x$Predicted.Output.probit)
```



```{r}
res <- list(Predicted.Output.logit=Predicted.Output.logit,
            Predicted.Output.probit=Predicted.Output.probit)
```

```{r}
write.table(res, file = paste(dataPath,'result.csv',sep = '/'), row.names = F)

```









