---
title: "LNL Assignment 8"
author: "Elyse Zhang"
date: "8/16/2018"
output: html_document
---

## 1 Example
### 1.1 Data

File MarketingExperiment.csv contains data reflecting total time spent by visitors of both genders on a web site and their probabilities of conversion.

Look at the sample in the file. Below nSample is the length of the sample imported from the file.


```{r}
dataPath = '/Users/Elyse/Documents/UChicago/Courses/Summer 2018/Linear Nonlinear/Course Material/Lecture 8'
```

```{r}
MarketingData<-read.csv(file=paste(dataPath,"MarketingExperiment.csv",sep="/"),header=TRUE,sep=",")
head(MarketingData)
```

```{r}
plot(MarketingData$Time,MarketingData$Probability, type="p",pch=19,xlab="Time",ylab="Probability")

```

### 1.2 Linear model
1.2.1 Probability as response time as predictor

Estimate linear model with one predictor Time using lm().
```{r}
lmProbabilityTime<-lm(Probability~Time,data=MarketingData)
lmProbabilityTime$coefficients
## (Intercept)        Time 
## 0.007165356 0.039232604
summary(lmProbabilityTime)
```

Analyze the residuals

```{r}
residuals<-lmProbabilityTime$residuals
plot(MarketingData$Time,residuals)
```

```{r}
density.Residuals<-density(residuals)
plot(density.Residuals,ylim=c(0,10))
lines(density.Residuals$x,dnorm(density.Residuals$x,mean=mean(residuals),sd=sd(residuals)))
```

```{r}
qqnorm(residuals)
qqline(residuals)
```

What do you conclude from the analysis of residuals?
From two population

1.2.2 Probability as response, time and gender as predictors
Add fixed effects based on gender.
Compare the two summaries.
```{r}
lmProbabilityTimeGender<-lm(Probability~Time+Gender,data=MarketingData)
summary(lmProbabilityTimeGender)
```

```{r}
summary(lmProbabilityTime)
```

1.2.3 Time as response, gender as predictor
Fit fixed effects model with time as predictor.

```{r}
lmTimeGender<-lm(Time~Gender,data=MarketingData)
summary(lmTimeGender)
```

```{r}
unique(predict(lmTimeGender))
```

Extract group means, call them FixedEffects
```{r}
Manual<-aggregate(MarketingData$Time,
                         by=list(MarketingData$Gender),
                         function(z) mean(z))$x
(Manual<-Manual-mean(Manual))
```

```{r}
fixedEffects<-c(lmTimeGender$coefficients[1],sum(lmTimeGender$coefficients))
(fixedEffects<-fixedEffects-mean(fixedEffects))

```

```{r}
cbind(Manual,fixedEffects)
```

### 1.3 Random effects
Learn how to fit the model using lmer() from library lme4

1.3.1 Time as response, gender as predictor
First, use the simplest random effects model with Time as output and Gender as random effect.

```{r}
suppressWarnings(library(lme4))
lmerTimeGender<-lmer(Time~1+(1|Gender),data=MarketingData)
summary(lmerTimeGender)
```

**What is the meaning of group random effect parameter Gender?**

Extract random effects.

```{r}
suppressWarnings(library(lattice))
(randomEffects<-ranef(lmerTimeGender,condVar=T))
```

```{r}
dotplot(randomEffects)
```

Check the meaning of random effects: deviations of random gender means from the base level.

```{r}
summary(lmerTimeGender)$coefficients
##             Estimate Std. Error  t value
## (Intercept) 13.02104 0.09790661 132.9945
baseLevel<-summary(lmerTimeGender)$coefficients[1]
pred<-unique(predict(lmerTimeGender))
pred<-sort(pred)
cbind(Manual=pred-baseLevel,
      RandomEffects=randomEffects$Gender)
```

Variance-correlation matrix returns standard deviations.
```{r}
suppressWarnings(library(arm))
VarCorr(lmerTimeGender)
##  Groups   Name        Std.Dev.
##  Gender   (Intercept) 0.080399
##  Residual             2.518331
sqrt(sum(se.ranef(lmerTimeGender)$Gender^2)/2+var(randomEffects$Gender))
##             (Intercept)
## (Intercept)  0.08036591
sd(residuals(lmerTimeGender))
```

Function se.ranef() from package arm returns standard errors of the model coefficients.

One way of thinking about the variances returned by the summary of lmer() is: residual variance is the variance within the groups and each random effect variance is the variance between the groups.

Compare standard deviation of random effects with standard deviation of fixed effects to measure shrinkage.

```{r}
(sdFixed<-sd(fixedEffects))
## [1] 0.1386281
(sdRandom<-sd(unlist(randomEffects)))
## [1] 0.04662787
sdRandom/sdFixed
```

1.3.2 Conversion probability as response, time and gender as predictors
Now apply lmer() to fit the model for probability with one predictor Time and one random effect based on Gender

```{r}
lmerProbabilityTimeGender = lmer(Probability~ 1+ Time +(1|Gender),data=MarketingData)
```

Extract fixed effects and random effects from the corresponding models.
```{r}
#Fixed
(lmCoeff<-lmProbabilityTimeGender$coefficients)
## (Intercept)        Time     GenderM 
##  0.04090073  0.03982643 -0.07722360
fixedEffects<-c(lmCoeff[1],lmCoeff[1]+lmCoeff[3])
(fixedEffects<-fixedEffects-mean(fixedEffects))
## (Intercept) (Intercept) 
##   0.0386118  -0.0386118
(randomEffects<-unlist(ranef(lmerProbabilityTimeGender)$Gender))
## (Intercept)1 (Intercept)2 
##   0.03859567  -0.03859567
```

Find coefficient of shrinkage as ratio of standard deviations.
```{r}
sd(randomEffects)/sd(fixedEffects)
```

**Which model fits the data better**
```{r}
plot(MarketingData$Time,predict(lmProbabilityTimeGender))
points(MarketingData$Time,predict(lmerProbabilityTimeGender),col="red")
```

```{r}
c(Fixed=AIC(lmProbabilityTimeGender),Random=AIC(lmerProbabilityTimeGender))

```


## Assignment
```{r}
dat <- read.table(paste(dataPath,'Week8_Test_Sample.csv',sep = '/'), header=TRUE)
```


```{r}
mFixedEffects = lm(dat$Response ~ dat$Predictor1 + dat$Predictor2)
summary(mFixedEffects)
```

```{r}
mRandomEffects = lmer(dat$Response ~ dat$Predictor1+ (1|dat$Predictor2))
summary(mRandomEffects)
#names(summary(mRandomEffects))
```

```{r}
(grand.mean<-mean(dat$Response))

(group.means = aggregate(dat$Response,by=list(dat$Predictor2),FUN=mean))
```



```{r}
fixedEffects.2<-c(mFixedEffects$coefficients[1],mFixedEffects$coefficients[1]+mFixedEffects$coefficients[-c(1)])

(fixedEffects.2 = fixedEffects.2[c(1,3:9)])

fixedEffects = c(0.6322970,1.2955929,1.4713489,2.8267232,1.2818048,1.1044656, 0.4003363, 1.6075606)
mean(fixedEffects)
(fixedEffects = fixedEffects-mean(fixedEffects))
```

```{r}
#(fixedEffects = (group.means-grand.mean)$x)
```


```{r}
(randomEffects = ranef(mRandomEffects))
#(randomEffects = randomEffects$`dat$Predictor2`$(Intercept))
(randomEffects = randomEffects[["dat$Predictor2"]][["(Intercept)"]])
```


```{r}
(fixedEffects/randomEffects)
```

```{r}
(Kappa = sd(randomEffects)/sd(fixedEffects))
```


```{r}
res <- list(fixedEffects = fixedEffects,
            randomEffects = randomEffects,
            Kappa = Kappa
            )
```

Save res to a file and upload the file using left sidebar.
```{r}
saveRDS(res, file = paste(dataPath,'result.rds',sep = '/'))
```







