---
title: "Assignment"
author: "Elyse Zhang"
date: "7/27/2018"
output: html_document
---
## 1 Measles immunization example
### 1.1 Data
Recent outbreak of measles caused debates about necessity of vaccination.

The data file for this project contains CDC data on immunization coverage for MMR for all U.S. states and the national immunization rate.

The data for outbreaks in each state were simulated using the assumption that after vaccination 90% of vaccinated children become immuned.

Another assumption was made that 100% of not immunized people get infected when exposed to the virus.

CDC data on immunization coverage rates

The data for this project are in the file MeaslesImmunizationCoverageAndOutbreaks.csv

Read the data.

```{r}
measles.data<-read.csv(file=paste(dataPath,"MeaslesImmunizationCoverageAndOutbreaks.csv",sep="/"))
#measles.data<-as.data.frame(measles.data)
head(measles.data)
```

### 1.2 Fit Poisson Regression
Fitting Poisson regression to the measles outbreaks data means fitting the model
ln(λi)=β0+β1xi,
where xi is the coverage variable and λi are intensities of the observed counts.

If Y people out of hundred got sick then the intensity of measles cases is λ=Y/100.
Then the model is ln(Y/100)=β0+β1x.
```{r}
measles.poisson.model = glm(Outbreaks ~ Coverage, family = poisson, data = measles.data)
summary(measles.poisson.model)
```


```{r}
measles.poisson.model$coefficients
measles.poisson.model$deviance
measles.poisson.model$df.residual
measles.poisson.model$aic
```

```{r}
rbind(measles.poisson.model$linear.predictors, log(measles.poisson.model$fitted.values), predict(measles.poisson.model))
```


```{r}
rbind(measles.data$Outbreaks,measles.poisson.model$fitted.values, exp(measles.poisson.model$linear.predictors),predict(measles.poisson.model, type = 'response'))
```

Interpretation of the model.

What if coverage changes by 1%?

Then the intensity of the Poisson distribution will have relative change
λ1/λ2=exp(β1Δx),
where Δx is the change in coverage.

Let Δx=0.01. Then the change of intensity measured in percentage points is
λnew/λold=exp(β0+β1(x+Δx)−β0−β1x)=exp(β1Δx).

According to the model 1% increase in immunization should reduce the intensity of measles cases by:
```{r}
(1-exp(measles.poisson.model$coef[2]*.01))*100
```

### 1.3 Fit Negative Binomial regression
Use package MASS to fit negative binomial regression.

```{r}
library(MASS)
```

Learn how to fit negative binomial regression using glm.nb() from MASS.
Fit model Negative.Binomial.Model.glm.nb.
Look at the summaries of both fits: Negative.Binomial.Model.glm.nb andmeasles.poisson.model and compare them.
Learn how to interpret the output of theta in Negative.Binomial.Model.glm.nb.
Describe your comparison of the two models in the report.

```{r}
Negative.Binomial.Model.glm.nb <- glm.nb(Outbreaks ~ Coverage, data = measles.data)
summary(Negative.Binomial.Model.glm.nb)
```

There is no overdispersion

```{r}
test_dat <- read.table(paste(dataPath,'Week5_Test_Sample.csv',sep = '/'), header=TRUE)
head(test_dat)
```

```{r}
glm.poisson.fit = glm(Output ~ Predictor, family = poisson, data = test_dat)
summary(glm.poisson.fit)
```


```{r}
glm.nb.fit = glm.nb(Output ~ Predictor, data = test_dat)
summary(glm.nb.fit)
```

```{r}
theta = glm.nb.fit$theta
```

```{r}
dispersiontest(glm.poisson.fit,alternative="two.sided")
names(dispersiontest(glm.poisson.fit,alternative="two.sided"))
dispersion.test.p.value = dispersiontest(glm.poisson.fit,alternative="two.sided")$p.value
```

We do not reject the Null that mean value equals variance, aka that the distribution is a possion

```{r}
cbind.data.frame(test_dat$Output,predict(glm.poisson.fit, type = 'response'), predict(glm.nb.fit,type ='response'))
plot(test_dat$Predictor,test_dat$Output, col = 'black')
points(test_dat$Predictor,predict(glm.poisson.fit, type = 'response'), col = 'red')
points(test_dat$Predictor,predict(glm.nb.fit,type ='response'), col = 'blue')

predicted.values = predict(glm.poisson.fit, type = 'response')
```



```{r}
res <- list(predicted.values=predicted.values,  
            dispersion.test.p.value=dispersion.test.p.value,
            theta = theta)
saveRDS(res, file = paste(dataPath,'result.rds',sep = '/'))
```
















