---
title: "Week1 workshop"
author: "Elyse Zhang"
date: "6/20/2018"
output: html_document
---


```{r}
Linear.Model.Data<-read.csv(file="/Users/Elyse/Documents/UChicago/Courses/Summer 2018/Linear Nonlinear/Course Material/Lecture 1/Week1ProjectData.csv",header=TRUE,sep=",")
Linear.Model.Data[1:10,]
n = length(Linear.Model.Data$Input)
```

```{r}
Linear.Model.Data.Frame<-as.data.frame(Linear.Model.Data)
```

### Fit Linear model Using lm() and glm()
Estimate the model using lm() and glm()
```{r}
Linear.Model.Data.lm<-lm(Output~Input,data=Linear.Model.Data.Frame)
Linear.Model.Data.glm<-glm(Output~Input,family=gaussian(link="identity"),data=Linear.Model.Data.Frame)
```

Check and compare the names of the objects elements in both fits.
```{r}
names(Linear.Model.Data.lm)
names(Linear.Model.Data.glm)
#Compare the summaries of outputs.

summary(Linear.Model.Data.lm)
summary(Linear.Model.Data.glm)
names(summary(Linear.Model.Data.lm))

```

### Compare the Model Objects.
Compare the following fields in the model objects.
Calculate the outputs of glm() using the fields of lm()

1. coefficients
2. residuals
3. fitted.values
4. linear.predictors
5. deviance
    Use deviance()
    Calculate deviance manually based on the definition given in the lecture
6. Akaike Information Criterion aic
    Obtain it by using AIC()
    Calculate it manually using the definition given in the lecture.
7. y
8. null.deviance
9. dispersion
10. The saturated model and the null model: log-likelihood and AIC

#### 1. Coefficients.
Come from both functions and have the same meaning and values.
```{r}
cbind(Output.LM=Linear.Model.Data.lm$coefficients,Output.GLM=Linear.Model.Data.glm$coefficients)

```

#### 2. Residuals.
Generalized linear model is not estimated by minimizing SSE.
So the residuals do not have to be the same and as we will see later, for distributions other than Gaussian they do differ.
However, in the case of Gaussian linear model maximization of log-likelihood is equivalent to minimization of SSE.
This should result in the same values for the residuals of lm() as those of glm().
Compare the two sets of residuals.

```{r}
cbind(Output.LM=Linear.Model.Data.lm$residuals,Output.GLM=Linear.Model.Data.glm$residuals)[1:10,]

```



#### AIC
```{r}
From.AIC.Function<-AIC(Linear.Model.Data.lm)
AIC.From.glm<-Linear.Model.Data.glm$aic
c(From.AIC.Function,AIC.From.glm)
```

```{r}
pi
```

```{r}
Manual.Log.Likelihood<- function(Linear.Model.Fit) {
  n = length(Linear.Model.Fit$fitted.values)
  sigma = summary(Linear.Model.Fit)$sigma
  return(-n*log(2*pi*sigma^2)/2-sum((Linear.Model.Fit$residuals)^2)/(2*(sigma^2)))
}
 
  #sigma summary%sigma
  #y sample
  #mu fitted values
  #yi-mui is residual
```

```{r}
(Log.Likelihood = Manual.Log.Likelihood(Linear.Model.Data.lm))
(AIC.Manual = -2 * Log.Likelihood) +2*3
```

#### Dispersion
```{r}
#1 Dispersion
summary(Linear.Model.Data.glm)$dispersion

#2
summary(Linear.Model.Data.lm)$sigma^2

#3
(var(Linear.Model.Data.lm$residuals))*n/(n-1)
(var(Linear.Model.Data.glm$residuals))*n/(n-1)

```

####AIC
```{r}
Linear.Model.Data.Null.lm<-lm(Output~1,data=Linear.Model.Data)
Null.Log.Likelihood.logLik<-logLik(Linear.Model.Data.Null.lm)
Null.AIC<-AIC(Linear.Model.Data.Null.lm)


Log.Likelihood.Null<-Manual.Log.Likelihood(Linear.Model.Data.Null.lm)
```

```{r}
c(Log.Likelihood.Null=Log.Likelihood.Null,
  Log.Likelihood=Log.Likelihood, #correct model
  Null.Log.Likelihood.logLik=Null.Log.Likelihood.logLik)
```








