---
title: "HW 7"
author: "Elyse Zhang"
date: "5/9/2018"
output: html_document
---
## 1 Fitting linear models
### 1.1 Read the data
```{r}
datapath <- "~/Documents/UChicago/Courses/Statistical Analysis/Assignments/Week 7/"
Regression.ANOVA.Data<-as.matrix(read.csv(file=paste(datapath,"DataForRegressionANOVA.csv",sep="/"),header=TRUE,sep=","))
Regression.ANOVA.Data<-read.csv(file=paste(datapath,"DataForRegressionANOVA.csv",sep="/"),header=TRUE,sep=",")
head(Regression.ANOVA.Data)
```

### 1.2 Fit linear models using: no inputs, only Input1, only Input2, both Input1 and Input2.
```{r}
fit.1<-lm(Output~1,data=Regression.ANOVA.Data) # only intercept
fit.1.2<-lm(Output~1+Input1,data=Regression.ANOVA.Data)
fit.1.3<-lm(Output~1+Input2,data=Regression.ANOVA.Data)
fit.1.2.3<-lm(Output~.,data=Regression.ANOVA.Data)
```

```{r}
summary(fit.1.2)
anova(fit.1.2)
```

Explain the following outputs and calculate them using only data and linear model fit (for example, Regression.ANOVA.Data and fit.1.2):

Df
```{r}
anova(fit.1.2)$Df
```

```{r}
k.1.2 =2
total.df = length(Regression.ANOVA.Data$Output)
(residual.df.1.2 = total.df -k)
```

Sum Sq
```{r}
anova(fit.1.2)$"Sum Sq"
```
```{r}
(SSM.1.2 = sum((fit.1.2$fitted.values-mean(fit.1.2$fitted.values))^2))
(SSE.1.2 = sum((Regression.ANOVA.Data$Output - fit.1.2$fitted.values)^2))
```
F value
```{r}
anova(fit.1.2)$"F value"[1]
```
```{r}
(F.stats.1.2 = SSM.1.2/1/(SSE.1.2/residual.df.1.2))
```

P-value Pr(>F)
```{r}
anova(fit.1.2)$"Pr(>F)"[1]
```
```{r}
1-pf(F.stats.1.2,1,residual.df.1.2)
```
**What does “<2.2e-16” mean in the output of anova()?**
**probablity of (Fstatistics = 813.4763, giving beta 1 =0 e.g. input 1 is not useful for the model) is less than 2.2e-16**

### 2.3
**What is H0 for F value in anova(fit.1.2) and for F-ststistic in summary(fit.1.2)?**
**beta1 = 0**

Obtain R2 from anova() and calculate it manually
```{r}
summary(fit.1.2)$r.squared
```
```{r}
(r.squared.1.2 = SSM.1.2/(SSM.1.2+SSE.1.2))
```



### 2.2 Compare summary(fit.1) and anova(fit.1)
```{r}
summary(fit.1)
anova(fit.1)
```

Note that anova(fit.1)$"Sum Sq" is the same as sum(fit.1$residuals^2)
```{r}
c(anova(fit.1)$"Sum Sq",sum(fit.1$residuals^2)) #SSE =SST
```

and the numbers of degrees of freedom are also the same
```{r}
c(anova(fit.1)$Df,fit.1$df.residual,summary(fit.1)$df[2])
```

**Why anova table does not show fields F value and Pr(>F)?**
```{r}
#because SSE = SST, no SSM
(SSM.1 = sum((fit.1$fitted.values-mean(fit.1$fitted.values))^2))
(SSE.1 = sum((Regression.ANOVA.Data$Output - fit.1$fitted.values)^2))
```

**because F value should be SSM/1/(SSE/residual.df), in this case, 0 since SSM is 0, and P(>F) is 100%**

2.4 Compare summary(fit.1.3) and anova(fit.1.3)

```{r}
summary(fit.1.3)
anova(fit.1.3)
```

**What do you conclude from the anova table?**
**Input 2 might not be that significant, as given beta2 = 0, probablity of Fstats > 1.6393 is 0.201. **

**Also the SSM is very small compared to SSE**

Compare F-value
```{r}
c(F.value=anova(fit.1.3)$"F value"[1],Df=anova(fit.1.3)$Df,P.value=anova(fit.1.3)$"Pr(>F)"[1])
##     F.value         Df1         Df2     P.value 
##   1.6393254   1.0000000 498.0000000   0.2010141
summary(fit.1.3)$fstatistic
```

### 2.5 Compare summary(fit.1.2.3) and anova(fit.1.2.3)
```{r}
summary(fit.1.2.3)

anova(fit.1.2.3)
```

## 3 Use anova() to compare nested linear models
Function anova() can be used to compare nested linear models

Compare the following 2 outputs.
```{r}
anova(fit.1.2,fit.1.2.3)
summary(fit.1.2.3)
```

**Did adding Input.2 change RSS in the anova table?** 
**adding Input.2 does not change RSS in the anova table too much, **

**What do you conclude from Pr(>|t|) in summary(fit.1.2.3) and Pr(>F) in anova(fit.1.3,fit.1.2.3)? **
**Input 1 is siginificant whereas input 2 is not.**

**Why anova(fit.1.3,fit.1.2.3) returns P-value of F-statistic, but summary(fit.1.2.3) returns Pr(>|t|) of t-statisic?**
**Anova test whether any of the betas are not zero, thus F statistics, summary of linear model test each of the betas whether they are zeros, this t statistics**

Compare the following 2 outputs.
```{r}
anova(fit.1,fit.1.2.3)
c(anova(fit.1,fit.1.2.3)$F[2],summary(fit.1.2.3)$fstatistic[1])
```
**Explain what is H0 for F-test in summary(fit.1.2.3)**
**beta1= beta 2 = 0 **

```{r}
test_dat <- read.table(paste(datapath,'Week7_Test_Sample.csv',sep = '/'), header=TRUE)
```

```{r}
t.fit.1<-lm(Output~1,data=test_dat) # only intercept
t.fit.1.2<-lm(Output~1+Input1,data=test_dat)
t.fit.1.3<-lm(Output~1+Input2,data=test_dat)
t.fit.1.2.3<-lm(Output~.,data=test_dat)
```

```{r}
anova(t.fit.1.2)
```

```{r}
anova(t.fit.1.3)
```

```{r}
anova(t.fit.1,t.fit.1.2.3)
```



















