---
title: "HW 8"
author: "Elyse Zhang"
date: "5/15/2018"
output: html_document
---
## 1 Coagulation data (Examples 7.3.1 - 7.3.4)
The data are in the package faraway and in the file coagulationdata.csv.
In this experiment four different diets were tried on animals. Then researchers measured their effects on blood coagulation time.
```{r}
suppressWarnings(library(faraway))
```

```{r}
coagulation
plot(coag~diet, data=coagulation, pch=19,col="black")
```

Evaluate visually mean values of each group and whether the differences between them are significant to check your intuition later.

Calculate summary statistics for each group.
Create the 2-column matrix of group means and sizes.
```{r}
summaryByGroup<-aggregate(coag~diet,data=coagulation,FUN=summary)
summaryByGroup$coag #returns 6 columns, 4th one is mean
summaryByGroup$diet

Sizes=aggregate(coag~diet,data=coagulation,FUN=length)
Sizes$diet
Sizes$coag

means<-cbind(Means=summaryByGroup$coag[,4],Sizes$coag)

rownames(means)<-as.character(summaryByGroup$diet)
means
```

You also can work with each group separately using subset().
For example,
```{r}
Group1.dietA<-subset(coagulation,coagulation$diet=="A")
Group1.dietA

summary(Group1.dietA)

mean(Group1.dietA[,1])
```

### 1.1 ANOVA for the data
Fit linear model for coag~diet.
Observe the summary, look and interpret results of fitted linear model and regression ANOVA.
```{r}
coag.model = lm(coagulation$coag ~ coagulation$diet)
modelSummary = summary(lm(coagulation$coag ~ coagulation$diet))
modelSummary$coefficients

modelSummary$df

c(modelSummary$sigma,modelSummary$r.squared)

modelSummary$fstatistic

modelANOVA = anova(lm(coagulation$coag ~ coagulation$diet))

modelANOVA
```

**1. If the formula is coag~diet then why are we getting estimates of the parameters dietB, dietC, and dietD?**
distance of their means to diet A

**2. Analyze statistical significance of all parameters based on p-values and standard errors.**
B C and intercept are significant. Not D

**3. Analize the values of parameters and interpret them. For example, What does the value of coefficient for dietC tell you?**
Difference between C and A is 7

**4. Analize the goodness of fit based on the determination coefficient, F-statistic.**
A good fit

**5. Analyze the residuals.**
SSM = 228, SSE = 112, SST = 340

Create matrix with dummy variable inputs for ANOVA.

**Why are we creating 3 input variables if we are given 4 groups?**
Because there is intercept A with all the TRUEs
```{r}
model.matrix(lm(coagulation$coag ~ coagulation$diet))
```


```{r}
coag<-coagulation
coag$x1<-coag$diet=="B"
coag$x2<-coag$diet=="C"
coag$x3<-coag$diet=="D"
coag
```

Fit full and null linear models for coag~x1+x2+x3 (all inputs) and coag~1 (intercept only).
Compare the two models using anova.
```{r}
coag.model.full<-lm(coag~x1+x2+x3, data=coag) #the same as linear model coag model
coag.model.null<-lm(coag~1,data=coag)
summary(coag.model.null)
summary(coag.model.full)
anova(coag.model.null,coag.model.full)
```

Compare and explain the output of anova(coag.model.null,coag.model.full) with the outputs of summary(coag.model) and  anova(coag.model).

**1. Compare degrees of freedom.**
**2. Compare the sums of squares.**
**3. compare the p-values.**

Null is not correct, didn't explain any variation


Calculate manually the sum of squares shown in ANOVA table.

In order to do that we need grand mean and the vector of full length of group means.

```{r}
grand.mean = mean(coag$coag)
grand.mean

A.mean= summaryByGroup$coag[1,4]
B.mean= summaryByGroup$coag[2,4]
C.mean= summaryByGroup$coag[3,4]
D.mean=summaryByGroup$coag[4,4]
group.mean = c(rep(A.mean,Sizes$coag[1]),rep(B.mean,Sizes$coag[2]),rep(C.mean,Sizes$coag[3]),rep(D.mean, Sizes$coag[4]))

group.mean

# Calculate SST,SSE and SSM. Observe decomposition of the variance representation SST.
SST= sum((coag$coag-grand.mean)^2)
SSE = sum((coag$coag[coag$diet=='A'] - A.mean)^2, (coag$coag[coag$diet=='B'] - B.mean)^2, (coag$coag[coag$diet=='C'] - C.mean)^2, (coag$coag[coag$diet=='D'] - D.mean)^2 )

SSM= SST- SSE

c(SST=SST,SSE=SSE,SSM=SSM)

```

```{r}
anova(coag.model)

anova(coag.model.null,coag.model.full)

```

## 2 Experiment plan
Check what experiment plan (basis) R uses in lm with model.matrix (coag.model)

Fit alternative model without intercept.
Check its experiment plan (basis).
```{r}
coag.altmodel<-lm(coag~diet-1,data=coagulation)
summary(coag.altmodel)

anova(coag.altmodel)

model.matrix(coag.altmodel)

```

```{r}
test_dat <- read.table(paste('~/Documents/UChicago/Courses/Statistical Analysis/Assignments/Week 8/','Week8_Test_Sample.csv',sep = '/'), header=TRUE)

```

```{r}
anova(lm(test_dat$Output ~ test_dat$Treatment))
summary(lm(test_dat$Output ~ test_dat$Treatment))
SSM = 126.64 #between
SST = 126.64 + 201.96
SSE = 201.96 #within
```
```{r}
anova(lm(test_dat$Output ~ test_dat$Treatment-1))
summary(lm(test_dat$Output ~ test_dat$Treatment-1))
SSMalt = 740.35
SSEalt = 201.95 #within
(SSTalt = SSMalt+SSEalt)

(SST.test = sum((test_dat$Output- 0)^2))
```


















