---
title: "HW2"
author: "Elyse Zhang"
date: "6/30/2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

### Question 1:
Load the visitors.rda dataset, make a time plot of your data and describe the main features of the series.
```{r}
load('visitors.rda')
head(visitors)
plot(visitors)
```

* **There is a clear upward trend.**
* **There is quite strong seasonality.**
* **The seasonal fluctuation seems to be increasing in magnitude over time.** 
* **There is no obvious cyclic pattern.**

### Question 2:
What is the appropriate Holt-Winters method for this data set (multiplicative / additive)? why?

**It should be multiplicative Holt-Winters method because the seasonal fluctuation seems to be increasing in magnitude over time**

### Question 3
Use the hw() function to forecast the next 15 months using Holt-Winters’ methods. Experiment with the following methods
```{r}
library('fpp')
```

* Linear trend with additive seasonality
```{r}
m_la = hw(visitors, seasonal = 'additive', h = 15)
plot(m_la)
```

* Linear trend with multiplicative seasonality
```{r}
m_lm = hw(visitors, seasonal = 'multiplicative', h = 15)
plot(m_lm)
```

* Exponential trend with multiplicative seasonality without damping
```{r}
m_em = hw(visitors, seasonal = 'multiplicative',  exponential = TRUE, h = 15)
plot(m_em)
```

* Exponential trend with multiplicative seasonality and damping
```{r}
m_emd = hw(visitors, seasonal = 'multiplicative',  damped = TRUE, exponential = TRUE, h = 15)
plot(m_emd)
```

* Linear trend with multiplicative seasonality and damping
```{r}
m_lmd = hw(visitors, seasonal = 'multiplicative', damped = TRUE, h = 15)
plot(m_lmd)
```

### Question 4:
Use the accuracy() function to compare the Root-Mean-Square-Error (RMSE) values of the forecasts from the various methods. Which do you prefer and why?

```{r}
cbind(m_la.RMSE = as.data.frame(accuracy(m_la))$RMSE, m_lm.RMSE = as.data.frame(accuracy(m_lm))$RMSE, m_em.RMSE = as.data.frame(accuracy(m_em))$RMSE, m_emd.RMSE = as.data.frame(accuracy(m_emd))$RMSE, m_lmd.RMSE = as.data.frame(accuracy(m_lmd))$RMSE)

```

**I prefer the model with linear trend with multiplicative seasonality and damping, because it has least Root-Mean-Square-Error. **

### Question 5:
a) Use the checkresiduals() function to check that the residuals from the best model look like white noise and provide a summary of the model’s smoothing parameters using the summary() function.

```{r}
checkresiduals(m_lmd)
```

* **Mean of the residuals seem to be very close to zero**
* **Residuals are reasonably close to normal distribution**
* **There are some occational correlations in the residuals series**

```{r}
summary(m_lmd)
```

Smoothing parameters:

* **alpha = 0.6668 for level meaning that this forecasting has quick response to changes**
* **beta  = 0.0043 for trend, meaning it has a more moderate trend**
* **gamma = 1e-04 for seasonal component, meaning smaller seasonality component**
* **phi   = 0.98 is the damping factor signaling that the trend line changes relatively rapidly over time.**
