---
title: "Assignment2"
author: "Elyse Zhang"
date: "4/20/2018"
output: html_document
---


##Part B: Manipulating, Categorizing, Sorting and Grouping & Summarizing Data

## QUESTION 1
Set up the connection
```{r}
library(RMySQL)
library(dplyr)
library(ggplot2)
library(DBI)
connection = dbConnect(MySQL(),user="root", password="", 
                dbname="sakila", host="localhost")
```


**a) Show the list of databases.**
```{r}
myQuery1a <- "SHOW DATABASES;"
dbGetQuery(connection, myQuery1a)
```


**b) Select sakila database.**
```{r}
myQuery1b <- "USE sakila;"
dbGetQuery(connection, myQuery1b)
```


**c) Show all tables in the sakila database.**
```{r}
myQuery1c <- "SHOW TABLES;"
dbGetQuery(connection, myQuery1c)
```


**d) Show each of the columns along with their data types for the actor table.**
```{r}
myQuery1d <- 
"SELECT
  COLUMN_NAME, DATA_TYPE
FROM
  INFORMATION_SCHEMA.COLUMNS
WHERE
  TABLE_NAME = 'actor';"

# alt: myQuery1d <- "DESCRIBE actor;"
dbGetQuery(connection, myQuery1d)
```


**e) Show the total number of records in the actor table.**
```{r}
myQuery1e <- 
"SELECT
  COUNT(*) AS total_record 
FROM 
  actor;"
dbGetQuery(connection, myQuery1e)
```


**f) What is the first name and last name of all the actors in the actor table ?**
```{r}
myQuery1f <- 
"SELECT
  first_name, last_name
FROM 
  actor;"
dbGetQuery(connection, myQuery1f)
```


**g) Insert your first name and middle initial ( in the last name column ) into the actors table.**
```{r}
myQuery1g <- 
"INSERT INTO actor (
   first_name,last_name
) VALUES (
   'ELYSE','Z');"
dbGetQuery(connection, myQuery1g)
```

To show it worked
```{r}
myQuery1g.show <- 
"SELECT 
   * 
FROM
   actor 
ORDER BY 
   actor_id DESC 
LIMIT 1;"
dbGetQuery(connection, myQuery1g.show)
```


**h) Update your middle initial with your last name in the actors table.**
```{r}
myQuery1h <- 
"UPDATE 
   actor
SET 
   last_name = 'ZHANG'
WHERE 
   first_name = 'ELYSE';"
dbGetQuery(connection, myQuery1h)
```


**i) Delete the record from the actor table where the first name matches your first name.**
```{r}
myQuery1h <- 
"DELETE FROM
   actor
WHERE 
   first_name = 'ELYSE';"
dbGetQuery(connection, myQuery1h)
```


**j) Create a table payment_type with the following specifications and appropriate data types**
Table Name : “Payment_type”
Primary Key: "payment_type_id”
Column: “Type”
Insert following rows in to the table: 1, “Credit Card” ; 2, “Cash”; 3, “Paypal” ; 4 , “Cheque”
```{r}
myQuery1j<- 
"CREATE TABLE payment_type (
  payment_type_id SMALLINT UNSIGNED NOT NULL AUTO_INCREMENT,
  type VARCHAR(45) NOT NULL,
  PRIMARY KEY  (payment_type_id)
)ENGINE=InnoDB DEFAULT CHARSET=utf8;"
dbGetQuery(connection, myQuery1j)

myQuery1j2<-
"INSERT INTO payment_type VALUES (1,'Credi Card'),
(2,'Cash'),
(3,'Paypal'),
(4,'Cheque');"
dbGetQuery(connection, myQuery1j2)
```


**k) Rename table payment_type to payment_types.**
```{r}
myQuery1k<- 
"ALTER TABLE payment_type RENAME payment_types;"
dbGetQuery(connection, myQuery1k)
```


**l) Drop the table payment_types.**
```{r}
myQuery1l<- 
"DROP TABLE payment_types;"
dbGetQuery(connection, myQuery1l)
```


## QUESTION 2
**a) List all the movies ( title & description ) that are rated PG-13 ?**
```{r}
myQuery2a<- 
"SELECT 
	  title, description
FROM
    film
WHERE
    rating = 'PG-13';"
dbGetQuery(connection, myQuery2a)
```

**b) List all movies that are either PG OR PG-13 using IN operator ?**
```{r}
myQuery2b<- 
"SELECT 
	  title, description
FROM
    film
WHERE
    rating IN ('PG-13', 'PG');"
dbGetQuery(connection, myQuery2b)
```

**c) Report all payments greater than and equal to 2$ and Less than equal to 7$ ? # Note : write 2 separate queries conditional operator and BETWEEN keyword**
```{r}
myQuery2c<- 
"SELECT 
	  payment_id, amount
FROM
    payment
WHERE
    amount BETWEEN 2 AND 7;
"
dbGetQuery(connection, myQuery2c)
```

**d) List all addresses that have phone number that contain digits 589 Note : write 3 different queries**
```{r}
myQuery2d1<- 
"SELECT 
	  address
FROM
    address
WHERE
    phone LIKE '%589%';"
dbGetQuery(connection, myQuery2d1)
```

**start with 140**
```{r}
myQuery2d2<- 
"SELECT 
	  address
FROM
    address
WHERE
    phone LIKE '140%';"
dbGetQuery(connection, myQuery2d2)
```

**end with 589**
```{r}
myQuery2d3<- 
"SELECT 
	  address
FROM
    address
WHERE
    phone LIKE '%589';"
dbGetQuery(connection, myQuery2d3)
```

**e) List all staff members ( first name, last name, email ) whose password is NULL?**
```{r}
myQuery2e<- 
"SELECT 
	  first_name, last_name, email
FROM
    staff
WHERE
    password IS NULL;"
dbGetQuery(connection, myQuery2e)
```


**f) Select all films that have title names like ZOO and rental duration greater than or equal to 4**
```{r}
myQuery2f<- 
"SELECT 
	  *
FROM
    film
WHERE
    title LIKE '%ZOO%'
AND 
    rental_duration >= 4;"
dbGetQuery(connection, myQuery2f)
```

**g) What is the cost of renting the movie ACADEMY DINOSAUR for 2 weeks ? # Note : use of column alias**
```{r}
# assuming rental_rate is for the rental duration listed
myQuery2g<- 
"SELECT 
	  rental_rate * 14 / rental_duration AS rental_cost
FROM
    film f
WHERE
    title = 'ACADEMY DINOSAUR';"
dbGetQuery(connection, myQuery2g)
```

**h) List all unique districts where the customers, staff, and stores are located # Note : check for NOT NULL values**
```{r}
myQuery2h<- 
"SELECT 
    district, COUNT(*)
FROM
    address
WHERE
     district IS NOT NULL
GROUP BY
    district
HAVING COUNT(*) = 1;"
dbGetQuery(connection, myQuery2h)
```

**i) List the top 10 newest customers across all stores**
```{r}
myQuery2i<- 
"SELECT 
    *
FROM
    customer
ORDER BY create_date DESC
LIMIT 10;"
dbGetQuery(connection, myQuery2i)
```

## Question 3
**a) Show total number of movies**
```{r}
myQuery3a<- 
"SELECT 
    COUNT(*) AS numofmovies
FROM
    film;"
dbGetQuery(connection, myQuery3a)
```

**b) What is the minimum payment received and max payment received across all transactions ?**
```{r}
myQuery3b<- 
"SELECT 
    MAX(amount),
    MIN(amount)
FROM
    payment;"
dbGetQuery(connection, myQuery3b)
```

**c) Number of customers that rented movies between Feb-2005 & May-2005 ( based on paymentDate ).**
```{r}
myQuery3c<- 
"SELECT
    COUNT(*)
FROM
    payment
WHERE
    payment_date BETWEEN '2005-02-01' AND '2005-05-31';"
dbGetQuery(connection, myQuery3c)
```

**d) List all movies where replacement_cost is greater than 15$ or rental_duration is between 6 & 10 days**
```{r}
myQuery3d<- 
"SELECT 
	  *
FROM
    film
WHERE
    replacement_cost > 15
AND 
    rental_duration BETWEEN 6 AND 10;"
dbGetQuery(connection, myQuery3d)
```

**e) What is the total amount spent by customers for movies in the year 2005 ?**
```{r}
myQuery3e<- 
"SELECT
    SUM(amount)
FROM
    payment
WHERE
    payment_date BETWEEN '2005-01-01' AND '2005-12-31';"
dbGetQuery(connection, myQuery3e)
```

**f) What is the average replacement cost across all movies ?**
```{r}
myQuery3f<- 
"SELECT 
	  AVG(replacement_cost)
FROM
    film;"
dbGetQuery(connection, myQuery3f)
```

**g) What is the standard deviation of rental rate across all movies ?**
```{r}
myQuery3g<- 
"SELECT 
	  STD(rental_rate)
FROM
    film;"
dbGetQuery(connection, myQuery3g)
```

**h) What is the midrange of the rental duration for all movies**
```{r}
myQuery3h<- 
"SELECT 
	  (MAX(rental_duration)+MIN(rental_duration))/2 AS durationmidrange
FROM
    film;"
dbGetQuery(connection, myQuery3h)
```

## QUESTION 4
**a) Customers sorted by first Name and last name in ascending order.**
```{r}
myQuery4a<-
"SELECT 
    first_name,
    last_name
FROM
    customer
ORDER BY first_name ASC, last_name ASC;"
dbGetQuery(connection, myQuery4a)
```

**b) Group distinct addresses by district.**
```{r}
myQuery4b<-
"SELECT 
    district, 
    address
FROM 
    address
GROUP BY 
    district,address;
"
dbGetQuery(connection, myQuery4b)
```

**c) Count of movies that are either G/NC-17/PG-13/PG/R grouped by rating.**
```{r}
myQuery4c<-
"SELECT 
    rating, COUNT(*) AS numberrating
FROM 
    film
GROUP BY 
    rating 
HAVING rating in ('G', 'NC-17', 'PG-13', 'PG', 'R');"
dbGetQuery(connection, myQuery4c)
```

**d) Number of addresses in each district.**
```{r}
myQuery4d<-
"SELECT 
    district, COUNT(*) AS numberaddress
FROM 
    address
GROUP BY 
    district ASC;"
dbGetQuery(connection, myQuery4d)
```

**e) Find the movies where rental rate is greater than 1$ and order result set by descending order.**
```{r}
myQuery4e<-
"SELECT 
    title
FROM 
    film
WHERE
    rental_rate > 1
ORDER BY title DESC;"
dbGetQuery(connection, myQuery4e)
```

**f) Top 2 movies that are rated R with the highest replacement cost ?**
```{r}
myQuery4f<-
"SELECT 
    title, rating, replacement_cost
FROM 
    film
WHERE
    rating = 'R'
ORDER BY
    replacement_cost DESC
LIMIT 2;"
dbGetQuery(connection, myQuery4f)
```

**g) Find the most frequently occurring (mode) rental rate across products.**
```{r}
myQuery4g<-
"SELECT 
    rental_rate, COUNT(*) AS frequentrate
FROM
    film
GROUP BY
    rental_rate
ORDER BY
    frequentrate
LIMIT 1;"
dbGetQuery(connection, myQuery4g)
```

**h) Find the top 2 movies with movie length greater than 50mins and which has commentaries as a special features.**
```{r}
myQuery4h<-
"SELECT
    title, length, special_features
FROM
    film
WHERE
    length > 50
AND 
    special_features LIKE '%commentaries%'
ORDER BY
    length DESC
LIMIT 2;"
dbGetQuery(connection, myQuery4h)
```

**i) List the years with more than 2 movies released.**
```{r}
myQuery4i<-
"SELECT
    release_year, COUNT(*) AS numfilm_released
FROM
    film
GROUP BY 
	release_year
Having COUNT(*) >2
ORDER BY
    COUNT(*) ASC;"
dbGetQuery(connection, myQuery4i)
```

## Part C: Combining Data, Nested Queries, Views and Indexes, Transforming Data

##Question 1
**a) List the actors (firstName, lastName) who acted in more then 25 movies. (Note: Also show the count of movies against each actor)**
```{r}
myQueryC1a ="SELECT
    a.first_name,
    a.last_name,
    count(fa.actor_id) AS numFilmActed
FROM
    actor a
JOIN
    film_actor fa
ON
    a.actor_id = fa.actor_id
GROUP BY
    fa.actor_id
HAVING
    count(fa.actor_id)> 25
ORDER BY
    count(fa.actor_id) DESC;"
dbGetQuery(connection, myQueryC1a)
```


**b) List the actors who have worked in the German language movies.**
```{r}
myQueryC1b ="SELECT
    a.first_name,
    a.last_name,
    f.title
FROM
    actor a 
JOIN
    film_actor fa
ON
    a.actor_id = fa.actor_id
JOIN  
    film f
ON  fa.film_id = f.film_id
WHERE
    f.language_id = 6;"
dbGetQuery(connection, myQueryC1b)
```

**c) List the actors who acted in horror movies. Note: Show the count of movies against each actor in the result set.**
```{r}
myQueryC1c ="SELECT
    a.first_name,  
    a.last_name,
    COUNT(a.actor_id) AS numHorrorMovies
FROM   
    actor a  
JOIN  
    film_actor fa  
    ON   a.actor_id = fa.actor_id  
JOIN   
    film f  
    ON  fa.film_id = f.film_id
JOIN
    film_category fc
    ON  fc.film_id = f.film_id
JOIN
    category c
    ON fc.category_id = c.category_id
WHERE 
    c.name = 'Horror'
GROUP BY
    a.actor_id;"
dbGetQuery(connection, myQueryC1c)
```

**d) List all customers who rented more than 3 horror movies.**
```{r}
myQueryC1d ="SELECT
    ct.first_name,  
    ct.last_name
FROM   
    customer ct
JOIN
    rental rt
    ON   ct.customer_id = rt.customer_id  
JOIN   
    inventory inv  
    ON  rt.inventory_id= inv.inventory_id
JOIN
    film f
    ON inv.film_id = f.film_id
JOIN
    film_category fc
    ON  fc.film_id = f.film_id
JOIN
    category c
    ON fc.category_id = c.category_id
WHERE 
    c.name = 'Horror'
GROUP BY
	ct.customer_id
HAVING
    COUNT(ct.customer_id)>3;"
dbGetQuery(connection, myQueryC1d)
```

**e) List all customers who rented the movie which starred SCARLETT BENING**
```{r}
myQueryC1e ="SELECT
    ct.first_name,  
    ct.last_name
FROM   
    customer ct
JOIN
    rental rt
    ON   ct.customer_id = rt.customer_id  
JOIN   
    inventory inv  
    ON  rt.inventory_id= inv.inventory_id
JOIN
    film f
    ON inv.film_id = f.film_id
JOIN
    film_actor fa
    ON  fa.film_id = f.film_id
JOIN
    actor a
    ON fa.actor_id = a.actor_id
WHERE 
    a.first_name = 'SCARLETT'
    AND
    a.last_name = 'BENING'
GROUP BY
	ct.customer_id;"
dbGetQuery(connection, myQueryC1e)
```


**f) Which customers residing at postal code 62703 rented movies that were Documentaries.**
```{r}
myQueryC1f ="SELECT
    ct.first_name,  
    ct.last_name
FROM   
    customer ct
JOIN
    address addr
    ON   ct.address_id = addr.address_id
JOIN
    rental rt
    ON   ct.customer_id = rt.customer_id  
JOIN   
    inventory inv  
    ON  rt.inventory_id= inv.inventory_id
JOIN
    film f
    ON inv.film_id = f.film_id
JOIN
    film_category fc
    ON  fc.film_id = f.film_id
JOIN
    category c
    ON fc.category_id = c.category_id
WHERE 
    c.name = 'Documentary'
    AND
    addr.postal_code = '62703'
GROUP BY
	ct.customer_id;"
dbGetQuery(connection, myQueryC1f)
```

**g) Find all the addresses where the second address line is not empty (i.e., contains some text), and return these second addresses sorted.**

```{r}
myQueryC1g ="SELECT
    address, address2
FROM 
    address
WHERE
    address2 <>''
ORDER BY address2 ASC;"
dbGetQuery(connection, myQueryC1g)
```

**h) How many films involve a “Crocodile” and a “Shark” based on film description ?**
```{r}
myQueryC1h ="SELECT
    count(*)
FROM 
    film
WHERE
    description LIKE '%Crocodile%' OR description LIKE '%Shark%';"
dbGetQuery(connection, myQueryC1h)
```

**i) List the actors who played in a film involving a “Crocodile” and a “Shark”, along with the release year of the movie, sorted by the actors’ last names.**
```{r}
myQueryC1i ="SELECT
    a.first_name,
    a.last_name,
    f.title,
    f.release_year
FROM
    actor a
JOIN
    film_actor fa
    ON fa.actor_id = a.actor_id
JOIN
    film f
    ON f.film_id = fa.film_id
WHERE
    f.description LIKE '%Crocodile%' OR f.description LIKE '%Shark%'
ORDER BY a.last_name;"
dbGetQuery(connection, myQueryC1i)
```

**j) Find all the film categories in which there are between 55 and 65 films. Return the names of categories and the number of films per category, sorted from highest to lowest by the number of films.**
```{r}
myQueryC1j ="SELECT
    c.name,
    COUNT(*)
FROM
	category c
JOIN
    film_category fc
    ON fc.category_id = c.category_id
GROUP BY 
    fc.category_id
HAVING
    COUNT(*) BETWEEN 55 AND 65
ORDER BY
    COUNT(*) DESC;"
dbGetQuery(connection, myQueryC1j)
```

**k) In which of the film categories is the average difference between the film replacement cost and the rental rate larger than 17$?**
```{r}
myQueryC1k ="SELECT
    c.name
FROM
    category c
WHERE
    c.category_id IN(
    SELECT
        fc.category_id
	FROM
        film_category fc
    JOIN
        film f
        ON f.film_id =fc.film_id
	GROUP BY fc.category_id
    HAVING
		AVG(f.replacement_cost)-AVG(f.rental_rate) > 17);"
dbGetQuery(connection, myQueryC1k)
```


**l) Many DVD stores produce a daily list of overdue rentals so that customers can be contacted and asked to return their overdue DVDs. To create such a list, search the rental table for films with a return date that is NULL and where the rental date is further in the past than the rental duration specified in the film table. If so, the film is overdue and we should produce the name of the film along with the customer name and phone number.**

**This is not doable, I listed people's name and email and movie name if their rental time is longer than rental duration**
```{r}
myQueryC1l ="SELECT
    f.title,
    ct.first_name,
    ct.last_name,
    addr.phone,
    DATEDIFF(rt.return_date,rt.rental_date),
    f.rental_duration
FROM
    customer ct
JOIN
    address addr
    ON   ct.address_id = addr.address_id
JOIN
    rental rt
    ON   ct.customer_id = rt.customer_id
JOIN   
    inventory inv  
    ON  rt.inventory_id= inv.inventory_id
JOIN
    film f
    ON inv.film_id = f.film_id
WHERE
    DATEDIFF(rt.return_date,rt.rental_date) > f.rental_duration;"
dbGetQuery(connection, myQueryC1l)
```

**m) Find the list of all customers and staff given a store id # Note : use a set operator, do not remove duplicates**
```{r}
myQueryC1m ="SELECT
    ct.first_name AS customerFirstName,
    ct.last_name AS customerLirstName,
    stf.first_name AS staffFirstName,
    stf.last_name AS staffLastName
FROM
    customer ct
JOIN
    store st
    ON ct.store_id = st.store_id
JOIN
    staff stf
    ON stf.store_id = st.store_id
WHERE
    st.store_id = 1;"
dbGetQuery(connection, myQueryC1m)
```

## Question 2
**a) List actors and customers whose first name is the same as the first name of the actor with ID 8.**
```{r}
myQueryC2a ="SELECT 
    a.actor_id AS ID,
    a.first_name AS FirstName,
    a.last_name AS LastName
FROM
    actor a
WHERE
    a.first_name = 
    (SELECT 
        first_name
	FROM 
        actor WHERE actor_id = 8)
UNION

SELECT
    ct.customer_id,
    ct.first_name,
    ct.last_name
FROM
	customer ct
WHERE
    ct.first_name = 
    (SELECT 
        first_name
	FROM 
        actor WHERE actor_id = 8);"
dbGetQuery(connection, myQueryC2a)
```

**b) List customers and payment amounts, with payments greater than average the payment amount**
```{r}
myQueryC2b ="
SELECT
    ct.customer_id,
    ct.first_name,
    ct.last_name,
    pm.amount
FROM
	customer ct
JOIN
    payment pm
    ON pm.customer_id = ct.customer_id
WHERE
    pm.amount >
    (SELECT
        AVG(amount)
	  FROM
        payment);"
dbGetQuery(connection, myQueryC2b)
```

**c) List customers who have rented movies atleast once # Note: use IN clause**
```{r}
myQueryC2c ="
SELECT
	first_name,
    last_name
FROM
    customer
    WHERE customer_id IN
    (SELECT
        customer_id
    FROM
        rental
	GROUP BY customer_id);"
dbGetQuery(connection, myQueryC2c)
```

**d) Find the floor of the maximum, minimum and average payment amount**
```{r}
myQueryC2d ="
SELECT
    FLOOR(MAX(amount)) AS Maximum,
	  FLOOR(MIN(amount)) AS Minimum,
	  FLOOR(AVG(amount)) AS Average
FROM 
	  payment;"
dbGetQuery(connection, myQueryC2d)
```

## Question 3
**a) Create a view called actors_portfolio which contains information about actors and films (including titles and category).**
```{r}
myQueryC3a ="
CREATE OR REPLACE VIEW actors_portfolio AS
    SELECT 
        a.*,
        f.title,
        c.name as Category
    FROM
        actor a
	JOIN 
        film_actor fa
    ON fa.actor_id = a.actor_id
    JOIN 
        film f
    ON f.film_id = fa.film_id
    JOIN
        film_category fc
    ON f.film_id = fc.film_id
    JOIN
		category c
	ON fc.category_id = c.category_id;"
dbGetQuery(connection, myQueryC3a)
```

**b) Describe the structure of the view and query the view to get information on the actor ADAM GRANT**
```{r}
myQueryC3b1 ="
Describe actors_portfolio;"
dbGetQuery(connection, myQueryC3b1)
```

```{r}
myQueryC3b2 ="
SELECT *
FROM
    actors_portfolio
WHERE
    first_name = 'ADAM' AND last_name = 'GRANT';"
dbGetQuery(connection, myQueryC3b2)
```

**c) Insert a new movie titled Data Hero in Sci-Fi Category starring ADAM GRANT**
```{r}
# myQueryC3c1 ="
# INSERT INTO actors_portfolio(actor_id, first_name,last_name,last_update)
# values(71,'ADAM','GRANT',NOW());"
# dbGetQuery(connection, myQueryC3c1) #cannot include the actor_id due to error
#Error in .local(conn, statement, ...) : could not run statement: Duplicate entry '71' for key 'PRIMARY'
```

```{r}
# myQueryC3c2 ="
# INSERT INTO actors_portfolio(title)
# values('Data Hero');"
# dbGetQuery(connection, myQueryC3c2)

#Error in .local(conn, statement, ...) : could not run statement: Field of view 'sakila.actors_portfolio' underlying table doesn't have a default value
```

```{r}
myQueryC3c3 ="
INSERT INTO actors_portfolio(category)
values('Sci-Fi');"
dbGetQuery(connection, myQueryC3c3)
```

## Question 4
**a) Extract the street number ( characters 1 through 4 ) from customer addressLine1**
```{r}
myQueryC4a ="
SELECT
    substring(address,1,4) AS streetNumber
FROM
    address
WHERE 
    address_id IN
    (SELECT
        address_id
	FROM
        customer);"
dbGetQuery(connection, myQueryC4a)
```

**b) Find out actors whose last name starts with character A, B or C.**
```{r}
myQueryC4b ="
SELECT
    first_name,
    last_name
FROM
    actor
WHERE
    last_name LIKE 'A%' OR last_name LIKE 'B%' OR last_name LIKE 'C%';"
dbGetQuery(connection, myQueryC4b)
```

**c) Find film titles that contains exactly 10 characters**
```{r}
myQueryC4c ="
SELECT
    title
FROM
    film
WHERE
    CHAR_LENGTH(title) = 10;"
dbGetQuery(connection, myQueryC4c)
```

**d) Format a payment_date using the following format e.g "22/1/2016"**
```{r}
myQueryC4d ="
SELECT
    DATE_FORMAT(DATE(payment_date), '%d/%m/%Y') AS '%d/%m/%Y'
FROM
    payment
WHERE payment_id = 1;"
dbGetQuery(connection, myQueryC4d)
```

**e) Find the number of days between two date values rental_date & return_date**
```{r}
myQueryC4e ="
SELECT
    DATEDIFF(return_date,rental_date)
FROM
    rental;"
dbGetQuery(connection, myQueryC4e)
```

## Question 5
**Provide eight additional queries and indicate the specific business use cases they address.**

**a) Find top 100 customer who rented the most often and email address, use case is marketing upsale opportunity**
```{r}
myQueryC5a ="
SELECT
    c.first_name,
    c.last_name,
    c.email,
    COUNT(rtl.customer_id)
FROM
    customer c
JOIN 
    rental rtl
    ON c.customer_id = rtl.customer_id
GROUP BY
    c.customer_id
ORDER BY COUNT(rtl.customer_id) DESC
LIMIT 100;"
dbGetQuery(connection, myQueryC5a)
```


**b) Find top 100 customer who pays the most and email address, use case is marketing upsale opportunity**
```{r}
myQueryC5b ="
SELECT
    c.first_name,
    c.last_name,
    SUM(amount)
FROM
    customer c
JOIN 
    payment pm
    ON c.customer_id = pm.customer_id
GROUP BY
    c.customer_id
ORDER BY SUM(amount) DESC
LIMIT 100;"
dbGetQuery(connection, myQueryC5b)
```

**c) Return customer who never rented, use case is digital marketing cost saving, more specifically, not to waste resource contact or email them in the short time**
```{r}
myQueryC5c ="
SELECT
    c.first_name,
    c.last_name,
    c.email,
    COUNT(rtl.customer_id)
FROM
    customer c
JOIN 
    rental rtl
    ON c.customer_id = rtl.customer_id
GROUP BY
    c.customer_id
HAVING COUNT(rtl.customer_id) = 1;"
dbGetQuery(connection, myQueryC5c)
```

**d) Count movies that are most rented. business case would be, get some more inventory and increase revenue**
```{r}
myQueryC5d ="
SELECT
    f.title,
    COUNT(inv.inventory_id)
FROM
    rental rt
JOIN   
    inventory inv  
    ON  rt.inventory_id= inv.inventory_id
JOIN
    film f
    ON inv.film_id = f.film_id

GROUP BY inv.film_id
ORDER BY COUNT(inv.inventory_id) DESC;"
dbGetQuery(connection, myQueryC5d)
```

**e) which store has more customer, so that we may want to expand it**
```{r}
myQueryC5e ="
SELECT
    store_id,
    COUNT(customer_id)
FROM 
    customer
GROUP BY 
    store_id;"
dbGetQuery(connection, myQueryC5e)
```

**f) which store has more income, so that we may want to expand it**
```{r}
myQueryC5f ="
SELECT
    staff_id, #staff and store id are the same fore this schema
    SUM(amount)
FROM 
    payment
GROUP BY 
    staff_id;"
dbGetQuery(connection, myQueryC5f)
```

**g) Also check the inventory, so that if the store that has more customers has less inventory, that should be corrected.**
```{r}
myQueryC5g ="
SELECT
    COUNT(inventory_id)
FROM
    inventory
GROUP BY store_id;"
dbGetQuery(connection, myQueryC5g)
```

**h) find who are the most popular actors so that store can prepare more inventorys**
```{r}
myQueryC5h ="
SELECT
    a.first_name,
    a.last_name,
    COUNT(rt.rental_id)
FROM
    rental rt
JOIN   
    inventory inv  
    ON  rt.inventory_id= inv.inventory_id
JOIN
    film f
    ON inv.film_id = f.film_id
JOIN
    film_actor fa
    ON f.film_id = fa.film_id
JOIN 
    actor a
    ON fa.actor_id = a.actor_id
GROUP BY
    fa.actor_id
ORDER BY COUNT(rt.rental_id) DESC
LIMIT 100;"
dbGetQuery(connection, myQueryC5h)
```



















